\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-04-07 k-planar.tex
minor
removed all the \notes, minor revisions

% \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
% \item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
% \item \citet{BDJM} use it to show that every $n$-vertex planar graph has local vertex ranking using $O(\log n/\log\log\log n)$ colours.
% % \item The result of \citet{DEJGMM21} implies that, for every integer
% $n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
% subgraph of $U_n$. This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.  The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
% aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
% \end{compactitem}

All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes. \citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of a surface with $h$ handles and $c$ crosscaps is $2h+c$. The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$. Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs,\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges. A class $\mathcal{G}$ of graphs is \defin{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$. A minor-closed class is \defin{proper} if it is not the class of all graphs. For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class. A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar. A 1-apex graph is \defin{apex}.  A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.} and \citet{DEMWW22} do so for bounded-degree graphs in any minor-closed class. See
% \cref{Generalisations} for more precise statements and see
\citep{DHJLW21} for a survey on this topic.

@@

\subsection{$k$-Planar Graphs}
% \label{k_planar_section}

@@

% These theorems have applications in diverse areas, which we explore in \cref{Applications}.

% This theorem has applications in diverse areas, including queue layouts  \citep{DJMMUW20}, non-repetitive colouring  \citep{dujmovic.esperet.ea:planar}, $p$-centered colouring  \citep{DFMS21}, and adjacency labelling \citep{DEJGMM21}, which we explore in \cref{Applications}. For example, we prove that $k$-planar graphs have bounded non-repetitive chromatic number (for fixed $k$). Prior to the recent work of \citet{dujmovic.esperet.ea:planar}, it was even open whether planar graphs have bounded non-repetitive chromatic number.
%
% \referee{1}{3. pages 2 – 4: The paragraph following Theorem 2 and Section 1.3
% (except for the fact that the apex assumption is necessary) is repetition
% of things discussed earlier on in the introduction.}
%
% \note{PM}{Kind of, but I suggest we keep that paragraph and section anyway. Justification: They're both short and they describe the exact bounds for bounded genus graphs, which you don't get from the one-paragraph description in Section~1.0.}
%
% \note{DW}{I suggest we write, ``This theorem has applications in diverse areas, which we explore in \cref{Applications}....''}

@@

% \referee{2}{Page 3. $P$ is used twice in the statement of Theorem 3.  It is probably
% safer (and necessary?) to use a different variable for the second
% occurrence.}

% \note{DW}{Use $\SS$ for shortcut system, okay?}
% \note{PM}{Ok.  I've implemented this change.  I was careful, but will still reread.}

Theorem~\ref{PlanarProduct}(b), \cref{ShortcutProduct}, and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$. Some further observations presented in \cref{k_planar_section} lead to the improved result.  \cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs. Some examples are explored in \cref{examples}.

% \note{DW}{I suggest we replace ``\cref{first_example,mid_example,last_example}''by ``\cref{examples}''.}

% Here is one example. The \defin{$k$-th power} of a graph $G$ is the graph $G^k$ with vertex set $V(G^k):=V(G)$, where $vw\in E(G^k)$ if and only if $\dist_G(v,w)\leq k$.\footnote{For a graph $G$ and two vertices $v,w\in V(G)$, $\dist_G(v,w)$  denotes the length of a shortest path, in $G$, with endpoints $v$ and $w$.  We define $\dist_G(v,w):=\infty$ if $v$ and $w$ are in different connected components of $G$.} If $G$ has maximum degree $\Delta$, then $G^k = G^\SS$ for some $(k,2k\Delta^{k})$-shortcut system $\SS$; see \cref{PowerShortcut}. Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} then imply:

% \begin{thm}
% \label{kPowerBasic}
% For every planar graph $G$ with maximum degree $\Delta$ and for every integer $k\geq 1$, $G^k$ is contained in $H\boxtimes P\boxtimes K_{6k^2(k^2+3)\Delta^{k}}$ for some graph $H$ of treewidth at most $\binom{k+3}{3}-1$ and some path $P$.
% \end{thm}

% \cref{Examples} presents further examples of graph classes that can be constructed using shortcut systems, including certain types of map graphs, string graphs, and $k$-nearest neighbour graphs. \cref{ShortcutProduct} implies product structure theorems for each of these classes. All of the above-mentioned applications also hold for these examples.

@@

% \note{DW}{I suggest the table headers are not in bold}

@@

% \note{DW}{why is the $2g+6$ commented out from the table?}

@@

A \defin{layering} of a graph $G$
% is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$
% \note{DW}{just write ``
is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge
 % ...''}
$vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.  For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$. To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the corresponding vertex in $H$.

@@

% \citet{DJMMUW20} introduced the study of partitions with bounded layered width such that the quotient has some additional desirable property, like small treewidth. Dujmovi\'c~et~al.\ define a class $\mathcal{G}$ of graphs to \defin{admit bounded layered partitions} if there exist $t,\ell\in\mathbb{N}$ such that every graph $G\in \mathcal{G}$ has an $H$-partition of layered width at most $\ell$ for some graph $H=H(G)$ of treewidth at most $t$.

@@

% \note{PM}{This feels out of place now that this section has been moved out of the introduction.  Delete?
%
% At the core of their work is the elegant proof by \citet{PS21} of the following result:
%
% \begin{thm}[\citep{PS21}]
% \label{ps}
%   Every planar triangulation $G$ has an $H$-partition $\PP$ such that $\tw(H)\leq 8$ and $G[P]$ is a shortest path in $G$ for each $P\in\PP$.
% \end{thm}
%
% Indeed, the above-mentioned result of \citet{DJMMUW20} is a slight strengthening of \cref{ps}, where for each $P\in\PP$ no two vertices of $P$ have the same distance to some fixed root vertex $r$.
% }

@@

% For each node $x\in V(T)$, define
% \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
% to be the maximal subtree of $T$ rooted at $x$.

% \note{PM}{I have significantly simplified this section, which should make the referees happy. I will read it very carefully again tomorrow}

@@

\begin{lem}\label{nice-decomposition}
  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a width-$t$ $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:
  \begin{compactenum}[(T1)]
    % \item\tlabel{rooted}\tlabel{first} $T$ is rooted at some node $x_0\in V(T)$ with $B_{x_0}=\emptyset$.
    % \item\tlabel{node-set} $V(T)= V(H)$.
    % \item\tlabel{diff} For each edge $xy\in E(T)$, $|B_x\ominus B_y|\le 1$.
    \item\tlabel{subtree-root} for each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$; and consequently
    \item\tlabel{ancestor-edge}\tlabel{last} for each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
  \end{compactenum}
\end{lem}

@@

% \referee{1}{6. page 6, last paragraph of Lemma 2: This isn’t quite right since you have to be careful with $x_0$.}
% \note{PM}{I addressed this by removing $x_0$ from $T_0$ at the beginning of the second paragraph.}

%\subsection{Generalized Tripod Partitions}
%
%
% \begin{figure}[htbp]
%   \begin{center}
%     \includegraphics{figs/tripoddo}
%   \end{center}
%   \caption{The sets $Y_x$, $F_x$, and $V_x$ associated with $x\in V(T)$
%   and the ancestors $a_1,\ldots,a_{t'}$ of $X$ such that $F_x \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.}
%   \label{fig:generalized-tripod}
% \end{figure}

% The following lemma shows how to interpret an $H$-partition of $G$ and a tree-decomposition of $H$ as a `hierarchical' decomposition of $G$; refer to \cref{fig:generalized-tripod}.

% For a vertex $v$ of $G$, $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[v]:=N_G(v)\cup\{v\}$.  For a vertex set $S\subseteq V$, $N_G(S):=\bigcup_{v\in S} N_G(S)\setminus S$ and $N_G[S]:=N_G(S)\cup S$.

% \note{PM}{
% Guide to old notation:
% \begin{compactitem}
%   \item $V_x := \bigcup_{y\in V(T_x)} Y_y = G_x$
%   \item $F_x:=\{w\in V(G): vw\in E(G), v\in V_x,\, w\not\in V_x\}= N_G(G_x)$
%   \item $N_x:=V_x\cup F_x=N_G[G_x]$
% \end{compactitem}
% }
%
% \begin{lem}\label{generalized-tripod}
%   Let $G$ be a graph; let $\mathcal{L}:=\langle L_1,\ldots,L_h\rangle$ be a layering of $G$; let $\mathcal{Y}:Y_x: x\in V(H))$ be an $H$-partition of $G$ of layered width at most $\ell$ with respect to $\mathcal{L}$ where $H$ has treewidth at most $t$; and let $\mathcal{T}:=(B_x:x\in V(T))$ be a tree-decomposition of $H$ satisfying the conditions of \cref{nice-decomposition}.
%   For each $x\in V(T)$, let $H_x:=V(T_x)$ and let $G_x:=\bigcup_{y\in H_x}Y_y$.
%
%   \begin{compactenum}[(Y1)]
%     % \item\ylabel{thickness} $\mathcal{Y}=(Y_x: x\in V(T))$ is a partition of $V(G)$ of layered width at most $\ell$ with respect to $\mathcal{L}$.
%     % \item\ylabel{separator} For each $x\in V(T)$, $N_G(\phi(T_x))$ separates $\phi(T_x)$ from $N
%     % there is no edge $vw\in E(G)$ with $v\in V_x$ and $w\in V(G)\setminus N_x$.
%
%     \item\ylabel{ancestor-edge} For each $x\in V(T)$, there is a set $\{a_1,\ldots,a_{t'}\}$ of $t'\le t$ strict $T$-ancestors of $x$ such that $N_G(G_x) \subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
%   \end{compactenum}
% \end{lem}
%
% \begin{proof}
%   Let $\{a_1,\ldots,a_{t'}\}:= B_x\setminus\{x\}$.  Then, by the definition of a normalized tree decomposition $N_H(H_x))\subseteq\{a_1,\ldots,a_{t'}\}$.  Therefore $N_G(G_x)\subseteq \bigcup_{i=1}^{t'} Y_{a_i}$.
% \end{proof}

% \referee{1}{1. Lemma 3: This lemma is really about ``normalised'' tree-decompositions, and says that for each $x \in V(T)$, the set $V(T_x)$ has at most $t$ neighbours in $H$. Call this property (T3) and remove (Y1) through (Y5), which should now be obvious.}
%
% \note{PM}{I've added to the \texttt{response.tex} file explaining some of this.  Each of (Y1), (Y3), (Y4), and (Y5) have a one-sentence justification. The justification for (Y2) is just definition unpacking and then using the observation that the referee recommends.}
%
% \note{PM}{I plan to do an experiment with the referee's suggested notation, in a separate git branch to see how it turns out.}
%
% Before proving \cref{generalized-tripod} we point out more properties that are immediately implied by it:
%
% \begin{compactenum}[(Y1)]\setcounter{enumi}{2}
%   \item\ylabel{y-subsets} $Y_x\subseteq V_x$ for every $x\in V(T)$. \note{PM}{$Y_x$, for $x\in V(T)$ is not defined!}
%   \item\ylabel{containment-i} $V_x\subseteq V_a$ for every $T$-ancestor $a$ of $x$.
%   \item\ylabel{containment-ii}$N_x\subseteq N_a$ for every $T$-ancestor $a$ of $x$.
% \end{compactenum}

% Property~\yref{y-subsets} follows from the fact that $V_x$ is the union of several sets, one of which is $Y_x$.  Property~\yref{containment-i} follows from the definition of $V_x$ and the fact that $V(T_x)\subseteq V(T_a)$. To show Property~\yref{containment-ii} first note that, by \yref{containment-i} it suffices to consider vertices $w\in F_x=N_x\setminus V_x$. By definition, every vertex $w\in F_x$ is adjacent, in $G$, to a vertex $v\in V_x$.  By \yref{containment-i}, $v\in V_a$, so $w$ is either in $V_a$ or $w$ satisfies the condition $vw\in E(G)$, $v\in V_a$, and $w\not\in V_a$, so $w\in F_a$.  In either case $w\in N_a=V_a\cup F_a$.  Note that none of \yref{y-subsets}--\yref{containment-ii} depends on \yref{ancestor-edge} (which is important, since \yref{containment-i} is used to establish \yref{ancestor-edge} in the following proof).
%
%
% \begin{proof}[Proof of \cref{generalized-tripod}]
%   % Property~\yref{thickness} follows immediately from the fact that $V(T)=V(H)$  and the fact that $\mathcal{Y}$ has layered width at most $\ell$ with respect to $\mathcal{L}$.
%   Property \yref{separator} is immediate from the definitions of $F_x$ and $N_x$.  In particular, $(N_x,V(G)\setminus V_x)$ is a separation of $G$ with $F_x=N_x\cap(V(G)\setminus V_x)$.
%
%   To establish Property~\yref{ancestor-edge}, consider some vertex $w\in F_x$.  Since $w\in F_x$, there exists an edge $vw\in E(G)$ with $v\in V_x$ and $w\not\in V_x$.  Since $v\in V_x$, $v\in Y_{x'}$ for some $T$-descendant $x'$ of $x$ (possibly $x=x'$). Since $\mathcal{Y}$ is a partition, $w\in Y_{a}$ for some $a\not\in V(T_x)$.  Since $vw\in E(G)$, we have $x'a\in E(H)$.  By \tref{ancestor-edge}, one of $a$ or $x'$ is a $T$-ancestor of the other. Since $w\in Y_a\subseteq V_a$ and $w\not\in V_x\supseteq V_{x'}$, \yref{containment-i} rules out the possibility that $x'$ is a $T$-ancestor of $a$. Therefore, $a$ is a $T$-ancestor of $x$ which is a $T$-ancestor of $x'$.  Let $z_0,\ldots,z_r$ be the path in $T$ from $z_0:=x'$ to $z_r:=a$.  Since $x'a\in E(H)$, at least one of $a$ or $x'$ is in $B_{z_i}$, for each $i\in\{0,\ldots,r\}$ and there is at least one $i$ such that $B_{z_i}$ contains both $x'$ and $a$.  However, by \tref{subtree-root} $x'$ is not contained in $B_{z_i}$ for any $i\in\{1,\ldots,r\}$.  Therefore $a\in B_{z_i}$ for each $i\in\{0,\ldots,r\}$.  In particular, $a$ is contained in $B_x$.
%   Property~\yref{ancestor-edge} now follows from the fact that $|B_x|\le t+1$ and $B_x$ contains $x$.
% \end{proof}

% \referee{2}{Page 7.  In the very last sentence, I think $B_{x_i}$ should be $B_{z_i}$
% (twice).  Moreover, the way the proof is written, it seems to suggest
% that we only get that $a \in B_{z_i}$ for each $i \in \{1, \dots, r\}$.
% The conclusion that $a \in B_{z_i}$ for each $i \in \{0, \dots, r\}$ is
% true though.  In a normalized tree-decomposition, if $xy$ is an edge and
% $x$ is a $T$-ancestor of $y$, then $x$ must be in $B_y$ (since the subtrees $T_x$
% and $T_y$ intersect).}

%\note{PM}{I addressed the preceding referee's.}

% We are now ready to complete the proof.
% , which we restate here for convenience:

% \mmg*

@@

% \note{DW}{These definitions should appear earlier, certainly before the definition of framed graph, which should be defined as an ``embedded graph''. One option is to put this whole paragraph as a footnote in the introduction where we first mention $k$-planar graphs. The experienced reader can skip this paragraph, which says to me it should be a footnote.}

@@

If an embedded graph $G$  has no crossings then it is \defin{non-crossing} graph.  When $G$ is non-crossing and $F$ is a face of $G$ then $V(F)$ denotes the set of vertices in $G$ on the boundary of $F$.  When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.  If  $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings, then $G$ is a $k$-plane graph.  A graph $G$ is \defin{$k$-planar} if is isomorphic to some $k$-plane graph.  A $0$-plane graph is a \defin{plane} graph and a $0$-planar graph is a \defin{planar} graph.

@@

% (This assumption is already present in the informal definition of $k$-planar graphs used in \cref{k_planar_section}.)

%
% \note{DW}{We need to assume that at most two edges cross at a single point, so we can add a dummy vertex of degree 4, right?}\note{PM}{No, under the current definition, $r$ edges crossing at a single point $p$ generate $\binom{r}{2}$ crossings, with each edge taking part in $r-1$ crossings, just like they would if we perturbed the edges so that each pair crosses at a distinct point.  I guess we could mention this.  (See the first paragraph in the proof of Theorem~2 below.)} \note{DW}{The downside of this approach is that in the opening section we give an informal definition of $k$-planar, which needs the assumption of having no three edges cross at a point, since we use this assumption in the construction of a $(k+1,2)$-shortcut system for $k$-planar graphs give in the Introduction. I think we should not have different definitions in two parts of the paper.} A (not necessarily embedded) graph $G'$ is \defin{$k$-planar} if there exists a $k$-plane graph $G$ isomorphic to $G'$. Under these definitions, $0$-planar graphs are exactly planar graphs and $0$-plane graphs are exactly plane graphs.

@@

%\begin{thm}
%\label{k-planar}
%Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $\tw(H)\leq \binom{k+4}{3}-1$.
%\end{thm}

\begin{proof}[Proof of \cref{kPlanarProduct}]
	Let $G$ be a $k$-plane graph.
    % We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$. This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.
	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$. In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.  Let $\SS := \{P_{vw}: vw\in E(G)\}$. Observe that $\SS$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\SS}\supseteq G$.  Specifically, $G_0^{\SS}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.

@@

	We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.  In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.  More precisely, each path $P\in\SS$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
	Redoing the calculation at the end of the proof of \cref{general-width} then yields
	\begin{align*}
	|S_x\cap Y_i| \le d\ell\left(
	2
	+ 4\left\lfloor\tfrac{k+1}{2}\right\rfloor
	+ 2\left\lceil\tfrac{k+1}{2}\right\rceil
	\right)
	 =
	d\ell\left(
	2 + 2(k+1) + 2\left\lfloor\tfrac{k+1}{2}\right\rfloor
	\right)
	 \le
	d\ell(3k+5)
	= 18k+30 \enspace .
	\end{align*}
	With this change, the layered width of the partition given by \cref{ShortcutPartition} becomes $(18k+30)(k+1)=18k^2+48k+30$.
	The result follows from \cref{PartitionProduct}.
	%This establishes \cref{k-planar}.
	% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.
\end{proof}

% \subsection{Examples}

@@

%\referee{2}{Page 16.  At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.  Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.  It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
%
%\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}

@@

% A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar.

\cref{StringShortcut,ShortcutProduct,GenusProduct} now imply:
 % and \cref{StringShortcut,qn} imply:

@@

% \begin{thm}
% \label{StringPartition}
% For integers $g\geq 0$ and $\delta\geq 2$, let $\ell:= \max\{2g,3\} \,(\delta^4 + 4 \delta^3 + 9 \delta^2 + 10 \delta + 4)$ and $t:= \binom{ \delta+4}{3}-1$.
% Then  every $(g,\delta)$-string graph:
% \begin{compactitem}
% 	\item is contained in $H\boxtimes P \boxtimes K_{\ell}$ for some path $P$ and for some graph $H$ with treewidth $t$,
% 	\item has queue-number $\qn(G^k) \leq 3 \ell \, 2^t - \ceil{\tfrac{3}{2}\ell}$.
% \end{compactitem}
% \end{thm}

% \note{DW}{\cref{StringPartition} has been updated.}

% Our results also give bounds on the non-repetitive chromatic number and the $p$-centered chromatic number of $(g,\delta)$-string graphs, but the bounds are weak, since such graphs $G$ have maximum degree at most $2\delta$, implying that $\pi(G) \leq (4+o(1))\delta^2$ and $\chi_p(G)\le p(64\delta)^2$ by results of \citet{DJKW16} and \citet{DFMS21}, respectively.

@@

% \note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}

@@

% \note{DW}{give a better version of \cref{MapPartition} with $g=0$. }

%So $(g,d)$-map graphs generalise graphs embedded in a surface, and we now assume that $d\geq 4$ for the remainder of this section.

@@

% \note{DW}{Suggested section name ``Graphs Derived From Framed Graphs'' or just``Framed Graphs''}

% Next we consider
%
% %%%%%%%%%%%%%%%%%%%
% \subsection{Framed Graphs}

% \note{PM}{I don't like this structure anymore.  The proof of \cref{kPlanarProduct} ($k$-planar) refers to the innards of the proof of \cref{ShortcutPartition} ($(k,d)$-shortcuts).  The proofs should be close to each other.
%
% Let's consider something like this:
% \begin{compactenum}
%   \item Introduction
%   \item Shortcut systems and $k$-planar graphs
%   \item All graphs obtained from shortcut systems or $k$-planar graphs
%   \item $d$-framed graphs, $1$-planar graphs, and map graphs
%   \item All applications (colouring, labelling, etc).
% \end{compactenum}
% % (Possibly with the order of 2 and 3 reversed.)
% }

% \note{DW}{I suggest we use the language of ``embedded graph'' here}
For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$. The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$. This definition (for simple graphs) was introduced by \citet{BDGGMR}.

@@

	%  \todo[inline]{%
	%    DW:Does this work for multigraph near-triangulations? \newline
	%    PM: Yes, it only requires that the inner faces have three sides and that the outer face is bounded by a simple cycle.  The proof goes through without change.  Look at the subgraph $X$ of the dual that contains only the edges dual to red/blue edges.  In our setting, the outer face has degree $1$ in $X$, which is odd.  Any bichromatic inner face has degree $0$ or degree $2$ in $X$.  There has to be another vertex of odd degree somewhere, which can only come from a trichromatic inner face.\newline
	%    DW: great. This is basically the same as the proof of the Hex Lemma.
	%  }

@@

% \note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}

>>>>

All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes. \citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of a surface with $h$ handles and $c$ crosscaps is $2h+c$. The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$. Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs,\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges. A class $\mathcal{G}$ of graphs is \defin{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$. A minor-closed class is \defin{proper} if it is not the class of all graphs. For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class. A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar. A 1-apex graph is \defin{apex}.  A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.} and \citet{DEMWW22} do so for bounded-degree graphs in any minor-closed class. See \citep{DHJLW21} for a survey on this topic.

@@

\subsection{$k$-Planar Graphs}

@@

@@

\cref{PlanarProduct}(b), \cref{ShortcutProduct}, and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$. Some further observations presented in \cref{k_planar_section} lead to the improved result.  \cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs. Some examples are explored in \cref{examples}.

@@

@@

@@

A \defin{layering} of a graph $G$ is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.  For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$. To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the corresponding vertex in $H$.

@@

@@

@@

@@

\begin{lem}\label{nice-decomposition}
  For every graph $H$ of treewidth $t$, there is a rooted tree $T$ with $V(T)=V(H)$ and a width-$t$ $T$-decomposition $(B_x:x\in V(T))$ of $H$ that has following additional properties:
  \begin{compactenum}[(T1)]
    \item\tlabel{subtree-root} for each node $x\in V(H)$, the subtree $T[x]:=T[\{y\in V(T):x\in B_y\}]$ is rooted at $x$; and consequently
    \item\tlabel{ancestor-edge}\tlabel{last} for each edge $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other.
  \end{compactenum}
\end{lem}

@@

@@

@@

If an embedded graph $G$  has no crossings then it is \defin{non-crossing} graph.  When $G$ is non-crossing and $F$ is a face of $G$, then let \defin{$V(F)$} be the set of vertices in $G$ on the boundary of $F$.  When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.  If  $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings, then $G$ is a $k$-plane graph.  A graph $G$ is \defin{$k$-planar} if is isomorphic to some $k$-plane graph.  A $0$-plane graph is a \defin{plane} graph and a $0$-planar graph is a \defin{planar} graph.

@@

@@

\begin{proof}[Proof of \cref{kPlanarProduct}]
	Let $G$ be a $k$-plane graph.	As in the proof of \cref{AddDummy}, let $G_0$ be the plane graph obtained by adding a dummy vertex at each crossing in $G$. In this way, each edge $vw\in E(G)$ corresponds naturally to a path $P_{vw}$ of length at most $k+1$ in $G_0$.  Let $\SS := \{P_{vw}: vw\in E(G)\}$. Observe that $\SS$ is a $(k+1,2)$-shortcut system for $G_0$ and that $G_0^{\SS}\supseteq G$.  Specifically, $G_0^{\SS}$ contains every edge and vertex of $G$ as well as the dummy vertices in $V(G_0)\setminus V(G)$ and their incident edges.

@@

	We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.  In this setting, the proof of \cref{general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $L_i\cap Y_x$.  More precisely, each path $P\in\SS$ containing an internal (dummy) vertex $w\in Y_x\cap (L_{i-j}\cup L_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap L_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap L_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap L_j$ for $j > k+1$.
	Redoing the calculation at the end of the proof of \cref{general-width} then yields
	\begin{align*}
	|S_x\cap Y_i| \le d\ell\left(
	2
	+ 4\left\lfloor\tfrac{k+1}{2}\right\rfloor
	+ 2\left\lceil\tfrac{k+1}{2}\right\rceil
	\right)
	 =
	d\ell\left(
	2 + 2(k+1) + 2\left\lfloor\tfrac{k+1}{2}\right\rfloor
	\right)
	 \le
	d\ell(3k+5)
	= 18k+30 \enspace .
	\end{align*}
	With this change, the layered width of the partition given by \cref{ShortcutPartition} becomes $(18k+30)(k+1)=18k^2+48k+30$.
	The result follows from \cref{PartitionProduct}.
\end{proof}

@@

@@

\cref{StringShortcut,ShortcutProduct,GenusProduct} now imply:

@@

@@

@@

@@

For any integer $d\ge 3$, an embedded (multi)graph $G$ is \defin{$d$-framed} if it has a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$. The embedded (multi)graph $G_0$ is called the \defin{frame} of $G$. This definition (for simple graphs) was introduced by \citet{BDGGMR}.

@@

@@
